---
title: "Day_08_WorkAlong"
author: "Analise Hofmann"
date: "April 4, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# While Loops & Logical Statements
While Loops:
  - web resource [1](http://www.endmemo.com/program/R/while.php)
  - web resource [2] (https://www.tutorialspoint.com/r/r_while_loop.htm)
Logicals:
  - web resource [1](http://www.astrostatistics.psu.edu/su07/R/html/base/html/Logic.html)

# Example 1
```{r}
# fix the below syntax, run the code, what happens?
# x <- 1
# while(x < 5) {x <- x+1; print(x);}
x<- 1
while( x < 5){
  x<- x+1
  print(x)
}

# how can we complete the same code above, but print 1 to 5, not just 2 to 5?
x<- 1
while( x <= 5){
  print(x)
  x<- x+1 #unlinke in a for loop we must manually increase the interation
}

# What happens if you change x <- 6
# while(x > 5) {x <- x-1; print(x);}
while (x>5){
  print(x)
  
  
  x<- x-1 #best to put interator at the end so that the value isn't changed until it gets to the next interation just in case the interator is used in the code.
}


```


### Activity 1
Write a while loop that prints out your name, one time for each letter of your name (aka. Analise needs to be printed 7 times)
  - useful website [1](http://stackoverflow.com/questions/1169456/the-difference-between-and-notations-for-accessing-the-elements-of-a-lis)
  - extra reading on [stringR] package (https://cran.r-project.org/web/packages/stringr/stringr.pdf)
```{r}
name1 <- "Analise"
# ?strsplit #use this to get help documentation
name2 <- strsplit(name1, split="")
name2[2] <- "Hofmann" #adding an element to the list
length(name2[[1]]) #gets us the length of the list element #1 (use double square brackets)
x<-1
while(x<=length(name2[[1]])){
  print(paste("interval is: ", x))
  # print(name1) #fix 1
   print(paste(name2[[1]], collapse = ""))
  x<- x+1
}

?paste #get help with



```


### Activity 2
Write a while loop that uses 'next'
```{r}
#convert the below code to normal syntax
# while(x < 5) { if (x == 3) next; print(x);x <- x+1;}
x<-1
y<-0
while(x<5){
  if(x==3){
    x<- x+1
    next
  }
  y <- y+x
  print(paste("iteration: ", x))
  print(paste("y is ",y))
  x<- x+1
}


````


### Activity 3
write a while loop that uses 'break' to exit the loop
```{r}
x<-1
y<-0
while(x<5){
  if(x==3){
    break
  }
  y <- y+x
  print(paste("iteration: ", x))
  print(paste("y is ",y))
  x<- x+1
}

```
