---
title: "Day 5, Loops Review & DplyR"
author: "Analise Hofmann"
date: "March 16, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Review 1

```{r}
#read in the table from Day 03
#practice changing working directories
day3table <-read.csv(file = "../Day_03_WorkAlong/TestData.csv")



```


## Review 2

```{r}
#use the table you read in above and write a for loop to modify it
day3table_original <- day3table
for(i in 1:3){ 
  #day3table[ 1, i ]<- 
  print (c(" column i is : ", i))  
  
  for ( j in 1:12){ 
    print(c("row j is: ", j))
      temp <- day3table[ j, i ] + 1 
     
      print (temp)
      day3table[ j, i ] <- temp
      
       }
  
}
```

## Connie's challenge to the group! 
Make a new column in day3table called "Sum" and sum the first three columns in each row, and put the sum into the new column
```{r}
day3table <- day3table_original #setting day3table back to original values

day3table[ ,"Sum"] <- NA # adding new empty column with NAs

#write a for loop to calculate the sum of each row and put it into our new column

for (i in 1:nrow(day3table)){
  
  temp <- sum(day3table[i, 1:3])
  # temp <- day3data[i,1] + day3data[i,2] + day3data[i,3] #Lucia's way
  
  # day3table[i, 4] <- temp # easiest way 
  day3table$Sum[i] <- temp #Lucia's way
  # day3table[i, "Sum"] <- temp #another alternate way
  
}






```
# Function to do sum of rows
```{r}
Sum3Column <- function(table){
  for (i in 1:nrow(table)){
    temp <- sum(table[i,1:3])
    table$Sum[i] <- temp
    }
  }

```

## Review 3
Challenge for homework before next meeting.
```{r}
#use the table above and write a while loop to modify it


```


## DplyR walk through from @JoeyBernhardt in UBC R Study Group
https://github.com/JoeyBernhardt/RStudyGroup-dplyr-lesson/blob/master/dplyr-lesson.Rmd

1. Let's look up the dplyR documentation. Where did you find it? What do you notice?
2. Let's start Joey's Lesson
3. Which libraries do we need? How can we download them? - google search

```{r}





```
