---
title: "Miniproject_Part2_Graphing"
author: "Analise Hofmann"
date: "April 18, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# get data
```{r}
library(dplyr)
library(ggplot2)

data_table_cnt <-read.csv(file = "./yeast-transcriptomics/data_table_normal.csv")

```

# Take a look at the file that was read in. 
 - what is happening? do we need to remove any columns?
```{r}
head(data_table_cnt)

#remove column 1
data_table_cnt <- data_table_cnt[,-1]
data_table_cnt <- data_table_cnt[,-4] # remove the ratio values for now

head(data_table_cnt)

```

# Do we want to normalize the experiments to a reference gene? How does that change the data? (sapply)
  - ACT1 (actin) ORF = YFL039C
  - TUB1 (alpha tubulin) ORF = YML085C
## get the raw values for each genotype for actin and tubulin  
```{r}
act1_row <- data_table_cnt[data_table_cnt$ORF == "YFL039C",]
tub1_row <- data_table_cnt[data_table_cnt$ORF == "YML085C",]

act1_WT <- act1_row[1,2]
act1_swr1 <- act1_row[1,3]

tub1_wt <- tub1_row[1,2]
tub1_swr1 <- tub1_row[1,3]

```

## normalize the data to actin
```{r}
data_act1 <- data_table_cnt

data_act1$WT <- sapply(data_table_cnt$WT, function(x) x/act1_WT)
data_act1$swr1 <- sapply(data_table_cnt$swr1, function(x) x/act1_swr1)

plot(data_act1$WT , data_act1$swr1, xlim = c(0,20), ylim= c(0,20))

plot(data_act1$WT , data_act1$swr1, xlim = c(0,4), ylim= c(0,4))

plot(data_act1$WT , data_act1$swr1, xlim = c(0,1), ylim= c(0,1))

#calculate the new ratios with a loop
for(i in 1:nrow(data_act1)){
  data_act1$Ratio[i] <- data_act1$WT[i] / data_act1$swr1[i]  
}



#easy way
data_act1$Ratio <- data_act1$WT/data_act1$swr1
data_act1$WT[5]/data_act1$swr1[5]
head(data_act1)
```

## normalize the data to tubulin
```{r}


```


# Convert the loop above into a function so that we can just call the function to calculate the ratios of tubulin normalized data.
BONUS: write the function so that it does the whole workflow above, aka also allows you to quickly choose a different refernce gene. 

```{r}
# gene name you want to normalize with
# data table you want to calculate
# column to normalize
#calculate the ratios for each row 





```

