---
title: "Miniproject Apr 11"
output: html_notebook
---
# Miniproject
For each of the following steps, let's make sure we understand what "objects" we're passing through to each step. Note: this miniproject is more about writing code, rather than good data analysis practices per se.

Data is from https://www.kaggle.com/costalaether/yeast-transcriptomics

## Preliminaries

Set working directory and read in CSV file.

```{r}
setwd("./Day_09_Miniproject/")

raw_data_table <- read.csv("yeast-transcriptomics/SC_expression.csv")
```

Inspect the data.

```{r}
head(raw_data_table, n=10)
tail(raw_data_table, n=10)
```

Looks like we have gene names and a bunch column labels. If we take a look at conditions_annotations.csv, we see that the different labels correspond to different experiments. Note: these data are generated from different labs with different conditions etc., but for the sake of this miniproject, we're going to pretend we can compare data directly.

Ok, let's take compare two experiments. AFIQCI corresponds to a WT experiment, AFNCCR corresponds to a swr1 mutant. We can use the $ operator to slice out the columns by their name.

```{r}
#Analise's Way, creates subsets of table
afi_afn_data <- raw_data_table[, c("X", "AFIQCI", "AFNCCR")]
#Andrew's Way, rips data to create more data
data_table <- data.frame(raw_data_table$X, raw_data_table$AFIQCI, raw_data_table$AFNCCR)
```
The colnames are hard to read, so let's rename the column names.
```{r}
colnames(data_table)
colnames(data_table) <- c("Gene", "WT", "swr1")
head(data_table)
```
Ok let's take a look at some quick summary statistics. Looks like we might not have a very well distributed data set.
```{r}
max(data_table$WT)
summary(data_table)
summary(data_table$WT)
```

Let's plot WT vs swr1 on x and y to see.

```{r}
#generic plotting
plot(data_table$WT, data_table$swr1)
#restricting max and min to get rid of outlier skewing graph
plot(data_table$WT, data_table$swr1, xlim=range(0,650), ylim=range(0,850))
```

Looks like there might be a massive outlier. Let's restrict the range to see more variation bottom left region.

```{r}
#structure
str(data_table)
#not the way to do it lol
data_table$WT > 1000
#true + true = 1
sum(c(TRUE, FALSE, TRUE))
#Therefore we can do this to get the number of data points above 1000
sum(data_table$WT > 1000)
```

How many values to we lose when we restrict the view like this?
```{r}

```

So either of these views we're still getting most of the data.

Since we're here to compare change between the two conditions, let's make a new column doing just that. We'll also log2 it, so that decreases become negatives, and increases become positives.


```{r}
#Add column of log2 of ratio between WT and swr1
data_table$log2ratios <- log2(data_table$WT/data_table$swr1)
#check
head(data_table)
#for removing a column....?
```

And for good measure, I'm going to create a new object where we sort this data_table. That way the genes at the top are the greatest increase, and the genes at the bottom are the greatest decrease. We can use the order function to perform this on data frames. order() returns a vector of indices in sorted order, which we can then use to call data_table rows in the appropriate order.
```{r}
#make new data table to sort the data, use command called 'order'
data_table_sorted <- data_table[order(data_table$log2ratios),]
#head table data is -inf b/c log of 0
head(data_table_sorted)
#tail table data is undefined b/c 0/0 is undefined
tail(data_table_sorted)
```

Uh oh. Looks like we get some weird values when either or both of the genes have 0 expression. Let's split up the data table into well behaved and not so well behaved subsets to see what's going on. We'll create a logical vector to find where WT or swr1 are == 0. The single pipe | operator is a vectorized inclusive OR. Because R coerces TRUE to 1 and FALSE to 0 when you try to add them, we can sum() to see how many abnormals we have

```{r}
data_table$WT == 0
data_table$swr1 == 0
test1 <- c(TRUE, FALSE)
test2 <- c(FALSE, TRUE)
test1 | test2
#
data_table_abnormals <- data_table_sorted$WT == 0 | data_table_sorted$swr1 == 0
#how many abnormal values are there?
sum(data_table_abnormals)
```

Now let's use this logical vector to split off our data set. We'll use ! to get the normal genes.
```{r}
# ! is called the "not" operator, gives you the opposite, in this case the FALSES
data_table_sorted_normal <- data_table_sorted [!data_table_abnormals,]
data_table_sorted_abnormal <- data_table_sorted [data_table_abnormals,]
head(data_table_sorted_normal)
head(data_table_sorted_abnormal)
```
Let's take a look at our normal data set now

```{r}

```

Let's take a look at the distribution. 

```{r}
#now we've excluded all the weird values of -inf and NaN, so can plot log2 values
plot(data_table_sorted_normal$log2ratios)
```