---
title: "Day 03 Work Along GitHub, Loops Etc"
author: "Analise Hofmann"
date: "February 15, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Intro to using [GitHub](www.github.com/ahofmann4/XX_CodingClub) 
**20-30 mins**

Today we will work through using GitHub online and on the desktop, practice pulling and pushing changes to our work. We will also discuss the pros of using GitHub for version control of coding progress, and how this can promote and faciliate collaborations on projects of all kinds (like this coding club! :) ). 

1. Walk through of online GitHub, (and GitHub Desktop if people want)
2. Integrating a GitHub repository with RStudio Project (pull this project to your computer)
3. Work in the repository and push & pull changes to the repository (make your own Notes_Name.Rmd file in Day_03 folder).

```{r, Read in file}
#ask everyone to read in the file using read.csv, do any problems come up?
#review checking the local directory, maybe you have to call the file differently?
#review setwd() and get() commands, '.' vs '..'

TestData <- read.csv(file = "./Day_03_WorkAlong/TestData.csv")

```

## What is an [if statement](https://www.programiz.com/r-programming/if-else-statement)?
** 15 mins **
- review boolean [logical statements](http://www.statmethods.net/management/operators.html)
- write if statements, and if else statements

```{r}


```

## What is a [for loop](http://paleocave.sciencesortof.com/2013/03/writing-a-for-loop-in-r/)?
** 20 mins **
- review a foor loop, and examples where you would use them
- nested for loops
- write a for loop to transfor a table
- can this be done easier with a specific function in R? apply? dplyR? 

```{r}


```

## **BONUS:** Write a function for your code above
