---
title: "if, else if, else, switch"
output: html_notebook
---

**Flow control** refers how our programs decide which parts of their code to execute. Often we'll want to execute code based on whether certain conditions are true. In programming, we use **conditionals** to do so. Conditionals are a basic skill of programming and must be mastered. For a quick reference, [see this link](https://www.tutorialspoint.com/r/r_decision_making.htm).

I'll illustrate with a real life example:

----

_If the weather is sunny outside, then I will go hiking._

_Or else if the weather is snowing, then I will go snowboarding._

_Otherwise, I'll work in the lab._

----

Notice the pattern: *If* ( some condition that is true or false ) { actions }.
In code, we formally refer to these components as: if ( boolean expression ) { statements }.

## Booleans
A quick review of Booleans. Boolean is a data type with two possible values: ```TRUE``` and ```FALSE```. We can construct various boolean expressions that perform logic and evaluate to ```TRUE``` or ```FALSE```, using [logical operators](http://www.statmethods.net/management/operators.html). 

```{r}
a <- TRUE
b <- FALSE
c <- 3
print(a && b)
print(a || b)
print (!a)
print(c==3)

```

## The **If** statement

The If statement executes a block of code only if it's conditional Boolean expression is ```TRUE```:

```{r}
a <- 3
if (a == 3) { #If a equals 3
  print("a is equal to 3")
}
```
The conditional is placed within parentheses, and the block of code within curly braces. If the expression is ```FALSE```, the code does not run.
```{r}
if (a != 3) { #If a does not equal 3
  print("a does not equal 3")
}
print("If there's no code above this line, the if statement did not execute")
```

Note: it is customary to indent code blocks within if statements. This makes for easier reading, and Rstudio should automatically be doing this for you.

You can nest if statements as much as you like:
```{r}
a <- 3
b <- 4
if (a == 3 ){
  if (b <= 5){
    print("a is 3 and b is smaller than 5")
  }
  if (b >= 5){
    print("a is 3 and b is smaller than 5")
  }
}
```
You can nest indefinitely, but then it gets ugly to read, and remember: good code tries to be easy to understand both by others, as well as yourself, including your future self, who may be only half an hour away.

Note: the boolean expression can be any that evaluates to TRUE/FALSE, so feel free to use variables in the conditional.
```{r}
a <- 3
b <- 6
c <- (a**2 + b/5) == 4
if (!c){
  print("wow c is false")
}
  
```

## If..else if.. else

When we want to decide between multiple choices, we can use the if...else if... else constructs:
```{r}
animal <- "dog"

if (animal == "cat"){
  print("Good kitty!")
} else if (animal =="dog"){
  print("Nice doggo!")
} else {
  print("UNRECOGNIZED MAMMAL ENCOUNTERED")
}

```
These conditionals are linked together in a construct (e.g. you can't have ```else if``` or ```else``` if you don't start with an ```if```. The code above first checks if the ```if``` condition is true, then if it's false, it will check if the ```else if``` condition is true. Finally, it will default to the code in the ```else``` block.

Note: a If...else if chain only will process UP TO a true condition:
```{r}
if(FALSE){
  print("1")
} else if (FALSE){
  print("2")
} else if (TRUE){
  print("3") # This executes
} else if (TRUE){
  print("4") # This does not.
}

```
Note: in the above code, I used if..else if, but I didn't use else, so there's no default option here.

## Switch()
If..else if... else will get you a lot of mileage. But sometimes it can be messy to write 20 else if statements. The ```switch()``` function can come in handy here. But it's a little more convoluted so we'll come back to this in another tutorial/problem set.

# Problems
## Problem 1 Warmup
Write a program that compares two integers and prints "equal" if they're equal, and "not equal" if they're not.
```{r}
a = 5 #change integers as required
b = 5
```
## Problem 2
Write a program that compares three integers and prints "equal" if they're ALL equal, and prints "not all equal" if they're not.

Bonus: Find the built-in R function that does this.
```{r}
# example integers
a = 5
b = 5
c = 6
```
## Problem 3
Given a logical vector (e.g. TRUE/FALSE), there is a built-in R function that will return TRUE if any of the elements are TRUE. Use your technical sophistication (e.g. your Google powers) to find this function.
```{r}
# Feeding this vector as an argument to this function will yield TRUE
vector_of_booleans = c(FALSE, FALSE, FALSE, TRUE, FALSE)
```
### Problem 4 Simple decision tree
Patients in a hospital can have symptom 1, or symptom 2, or both. In the following code, patients are given a two element vector, their status of having symptom 1 and 2 indicated as TRUE/FALSE in the 1st and 2nd element respectively. Separate patients into groups A for symptom 1, B for symptom 2, and C if they have both symptoms. Do this by changing patient_to_test, and then printing e.g. "Group A" depending on their symptoms. If they have no symptoms, print "No symptoms". Hint: you can accomplish this using nested conditionals, or using logical operators. Be sure to test your code on all 4 patients to see if it works.

```{r}
PATIENT_1 <- c(TRUE, FALSE)
PATIENT_2 <- C(FALSE, TRUE)
PATIENT_3 <- c(TRUE, TRUE)
PATIENT_4 <- c(FALSE, FALSE)
patient_to_test <- PATIENT_1 # change this to the patient you want to test.

# Your code goes here.


```
### Problem 5:
Pneumonoultramicroscopicsilicovolcanoconiosis is the longest word in English dictionaries. It is broken up in a vector of letters below. In the for loop provided, every time the letter "o" comes up, print "Letter O", and every time the letter "l" comes up, print "Letter L".

BONUS: Initialize some variables outside the for loop, and use if statements to count how many times the letter "o" comes up.
BONUS 2: Find another way of counting o's in the word without using the for loop at all!

```{r}
letters = strsplit("pneumonoultramicroscopicsilicovolcanoconiosis", split="")[[1]]

for (letter in letters){
  #Your code here
}

```


### Problem 6: FizzBuzz
The [FizzBuzz](https://en.wikipedia.org/wiki/Fizz_buzz) problem is an often-asked basic coding interview question (and a children's learning game). Given a number x between 1 and 100, if it's a multiple of 3, print "Fizz", if it's a multiple of 5, print "Buzz", if it's a multiple of both 3 and 5, print "FizzBuzz", otherwise just print the number. You can either assign a variable to be your "test number", or better yet, write your conditionals inside a for loop that iterates over a vector of integers from 1 to 100.
```{r}
# Option A
x <- 3
## You can change x to be the following: 3, 5, 12, or 15
## When x is 3, your conditionals should print "Fizz"
## When x is 5, your conditionals should print "Buzz"
## When x is 12, your conditionals should print "14"
## When x is 15, your conditionals should print "FizzBuzz".
## Note: Of course, you may want to check other numbers too, e.g. 20,21
## Hint: You should be able to do this without using the paste() function.

# Option B
## First write a for loop that iterates from 1 to 100.
## Put your conditional code inside the for loop.

```



